---
title: "4: Data Manipulation"
description: |
  Code for Quiz 5. More practice with dplyr functions.
date: 02-20-2021
output:
  distill::distill_article:
    self_contained: false
preview: preview.png        
---

1.  Load the R packages we will use.

```{r}
library(tidyverse)
```

2.  Read the data in the file, `drug_cos.csv` in to R and assign it to `drug_cos`.

```{r}
drug_cos  <- read_csv("https://estanny.com/static/week5/drug_cos.csv")
```

3.  Use `glimpse()` to get a glimpse of your data.

```{r}
glimpse(drug_cos)
```

4.  Use `distinct()` to subset distinct rows.

```{r}
drug_cos  %>% 
  distinct(year)
```

5.  Use `count()` to count observations by group.

```{r}
drug_cos  %>% 
  count(year)

drug_cos  %>% 
  count(name)

drug_cos  %>% 
  count(ticker, name)
```

## Use `filter()` to extract rows that meet criteria

6.  Extract rows in non-consecutive years

```{r}
drug_cos  %>% 
  filter(year %in% c(2013, 2018))
```

7.  Extract every other year from 2012 to 2018

```{r}
drug_cos  %>% 
  filter(year %in% seq(2012, 2018, by = 2))
```

8.  Extract the tickers "PFE" and "MYL"

```{r}
drug_cos  %>% 
  filter(ticker %in% c("PFE", "MYL"))
```

## Use `select()` to select, rename and reorder columns

9.  Select columns `ticker`, `name` and `ros`

```{r}
drug_cos  %>% 
  select(ticker, name, ros)
```

10. Use `select` to exclude columns `ticker`, `name` and `ros`

```{r}
drug_cos  %>% 
  select(-ticker, -name, -ros)
```

11. Rename and reorder columns with `select`

-   start with `drug_cos` THEN

-   change the name of `location` to `headquarter`

-   put the columns in this order: `year`, `ticker`, `headquarter`, `netmargin`, `roe`

```{r}
drug_cos  %>% 
  select(year, ticker, headquarter =location, netmargin, roe )

```

### Question: filter and select

Use inputs from your quiz question **filter and select** and replace **SEE QUIZ** with inputs from your quiz and replace the **???** in the code

-   start with `drug_cos` THEN
-   extract information for the tickers **PFE, MRK, BMY** THEN
-   select the variables `ticker`, `year` and `ros`

```{r}
drug_cos  %>% 
  filter(ticker %in% c("PFE", "MRK", "BMY"))  %>% 
  select(ticker, year, ros)
```


### Question: rename

- start with `drug_cos` THEN 
- extract information for the tickers  AGN, ZTS  THEN 
- select the variables `ticker`, `netmargin` and `roe`.  Change the name of `roe` to `return_on_equity`


```{r eval = FALSE}
drug_cos  %>% 
  ???(??? %in% c("???", "???"))  %>% 
  ???(ticker, ???, ??? = roe)
```

12. `select` ranges of columns

-  by name

```{r}
drug_cos  %>% 
  select(ebitdamargin:netmargin)
```


- by position

```{r}
drug_cos  %>% 
  select(4:6)

```

13. `select` helper functions

- `starts_with("abc")` matches columns start with "abc"

- `ends_with("abc")` matches columns end with "abc"

- `contains("abc")` matches columns contain "abc"


```{r}
drug_cos  %>% 
  select(ticker, contains("locat"))

drug_cos  %>% 
  select(ticker, starts_with("r"))

drug_cos  %>% 
  select(year, ends_with("margin"))
```


## Use `group_by` to set up data for operations by group

14. `group_by` 

```{r}
drug_cos  %>% 
  group_by(ticker)

drug_cos  %>% 
  group_by(year)

```


## Use `summarize` to calculate summary statistics  


15. Maximum `roe` for all companies

```{r}
drug_cos  %>% 
    summarize( max_roe = max(roe)) 
```

- maximum `roe` for each `year`


```{r}
drug_cos  %>% 
  group_by(year)  %>% 
  summarize( max_roe = max(roe)) 
```

- maximum `roe` for each `ticker`


```{r}
drug_cos  %>% 
  group_by(ticker)  %>% 
  summarize( max_roe = max(roe)) 
```


### Question: **summarize**

**Mean for year**

* Find the mean **SEE QUIZ** for each `year` and call the variable **mean_SEE QUIZ**

* Extract the mean for **SEE QUIZ** 


```{r eval = FALSE}
drug_cos  %>% 
  ???(???)  %>% 
  ???(??? = ???(???))  %>% 
  filter( year == ???)
```

* The mean ros for 2016 is  ??? or ???%


**Median for year**

* Find the median **SEE QUIZ** for each `year` and call the variable **median_SEE QUIZ**

* Extract the median for **SEE QUIZ** 


```{r eval = FALSE}
drug_cos  %>% 
  ???(???)  %>% 
  ???(??? = ???(???))  %>% 
  filter(year == ???)
```

* The median **SEE QUIZ** for **SEE QUIZ** is:  ??? or ???%

16. Pick a ratio and a year and compare the companies. 

```{r}
drug_cos  %>% 
 filter(year == 2018)  %>% 
 ggplot(aes(x = netmargin, y = reorder(name, netmargin))) +
 geom_col() +
  scale_x_continuous(labels = scales::percent) +
  labs(title = "Comparision of net margin",
       subtitle = "for drug companies during 2018", 
       x = NULL, y = NULL) +
  theme_classic() 
``` 

17. Pick a company and a ratio and compare the ratio over time. 

```{r}
drug_cos  %>% 
 filter(ticker == "PFE")  %>% 
 ggplot(aes(x = year, y = netmargin)) +
 geom_col() +
  scale_y_continuous(labels = scales::percent) +
  labs(title = "Comparision of net margin",
       subtitle = "for Pfizer from 2011 to 2018", 
       x = NULL, y = NULL) +
  theme_classic() 

ggsave(filename = "preview.png", 
       path = here::here("_posts", "2021-02-20-data-manipulation"))

``` 


