---
title: "5: Joining data"
description: |
  Code for Quiz 6, more dplyr and our first interactive chart using echarts4r.
date: 02-27-2021
output:
  distill::distill_article:
    self_contained: false
    toc: true
preview: preview.png   
---

# Steps 1-6 

1.  Load the R packages we will use.

```{r}
library(tidyverse)
library(echarts4r)  #install this package before using
library(hrbrthemes) #install this package before using
```

2.  Read the data in the files, `drug_cos.csv`, `health_cos.csv` in to R and assign to the variables `drug_cos` and `health_cos`, respectively 

```{r}
drug_cos  <- read_csv("https://estanny.com/static/week6/drug_cos.csv")
health_cos  <- read_csv("https://estanny.com/static/week6/health_cos.csv")
```

3. Use `glimpse` to get a glimpse of the data 

```{r}
drug_cos %>% glimpse()
health_cos %>% glimpse()  

```

4. Which variables are the same in both data sets

```{r}
names_drug  <- drug_cos  %>%  names() 
names_health  <- health_cos  %>%  names() 
intersect(names_drug, names_health)
```


5. Select subset of variables to work with

- For `drug_cos` select (in this order): `ticker`, `year`, `grossmargin`

  - Extract observations for 2018

  - Assign output to `drug_subset`

- For `health_cos` select (in this order): `ticker`, `year`, `revenue`, `gp`, `industry`

  - Extract observations for 2018

  - Assign output to `health_subset`




```{r}
drug_subset  <- drug_cos  %>% 
  select(ticker, year, grossmargin)  %>% 
  filter(year == 2018)

health_subset  <- health_cos  %>%
  select(ticker, year, revenue, gp, industry)  %>% 
  filter(year == 2018)

```

6. Keep all the rows and columns `drug_subset` join with columns in  `health_subset` 

```{r}
drug_subset  %>% left_join(health_subset)

```


# Question: join_ticker

* Start with `drug_cos` 

* Extract observations for the ticker **SEE QUIZ** from `drug_cos`
* Assign output to the variable `drug_cos_subset`

```{r eval = FALSE}
drug_cos_subset  <- drug_cos  %>% 
  ???(??? == "???")

```

*******************


* Display `drug_cos_subset`

```{r eval = FALSE}
drug_cos_subset

```


* Use left_join to combine the rows and columns of `drug_cos_subset` with the columns of  `health_cos`  

* Assign the output to `combo_df`

```{r eval = FALSE}

???  <- drug_cos_subset  %>% 
  left_join(health_cos)
 

```

*******************

* Display `combo_df`

```{r eval = FALSE}
combo_df

```

* Note: the variables `ticker`, `name`, `location` and `industry` are the same for all the observations

*******************

* Assign the company name to `co_name`

```{r eval = FALSE}

???  <- combo_df  %>% 
  distinct(name) %>% 
  pull()

```

*******************

* Assign the company location to `co_location`


```{r eval = FALSE}
co_location  <- ???  %>% 
  ???(???)  %>% 
  pull() 

``` 

*******************

* Assign the industry to `co_industry` group

```{r eval = FALSE}
co_industry  <- ???  %>% 
  ???(???)  %>% 
  ???() 

```

Put the r inline commands used in the blanks below.  When you knit the document the results of the commands will be displayed in your text.  

The company ??? is located in ??? and is a member of the ??? industry group.


*******************

* Start with `combo_df`

* Select variables (in this order): `year`, `grossmargin`, `netmargin`, `revenue`,  `gp`, `netincome`  

* Assign the output to `combo_df_subset`

```{r eval = FALSE}
combo_df_subset  <- combo_df  %>% 
  select(???, ???, ???, 
  ???, ???, ???)

```

*******************

* Display `combo_df_subset`

```{r eval = FALSE}

???

```

*******************

* Create the variable `grossmargin_check` to compare with the variable `grossmargin`.  They should be equal.
    - `grossmargin_check` = `gp` / `revenue`

* Create the variable `close_enough` to check that the absolute value of the difference between `grossmargin_check` and `grossmargin` is less than 0.001


```{r eval = FALSE}

combo_df_subset  %>% 
  mutate(grossmargin_check = ??? / ???,
  close_enough = abs(grossmargin_check - grossmargin) < 0.001)

```

*******************

* Create the variable `netmargin_check` to compare with the variable `netmargin`. They should be equal.

* Create the variable `close_enough` to check that the absolute value of the difference between `netmargin_check` and `netmargin` is less than 0.001


```{r eval = FALSE}
???  %>% 
  ???(netmargin_check = ??? / ???,
  close_enough = ???(netmargin_check - netmargin) < 0.001)
```


# Question: summarize_industry

* Fill in the blanks  

* Put the command you use in the Rchunks in the Rmd file for this quiz 

* Use the `health_cos` data

* For each industry calculate  
  * SEE QUIZ = mean(SEE QUIZ / revenue) * 100
  * SEE QUIZ = median(SEE QUIZ / revenue) * 100
  * SEE QUIZ = min(SEE QUIZ / revenue) * 100
  * SEE QUIZ = max(SEE QUIZ / revenue) * 100


```{r eval = FALSE}
health_cos  %>% 
  group_by(???)  %>% 
  summarize(??? = ???(??? / revenue) * 100,
            ??? = ???(??? / revenue) * 100,
            ??? = ???(??? / revenue) * 100,
            ??? = ???(??? / revenue) * 100
  ) 

```

* **SEE QUIZ** for the industry **SEE QUIZ** is ???%
* **SEE QUIZ** for the industry **SEE QUIZ** is ???%
* **SEE QUIZ** for the industry **SEE QUIZ** is ???%
* **SEE QUIZ** for the industry **SEE QUIZ** is ???%


# Question: inline_ticker

* Fill in the blanks  
 
* Use the `health_cos` data

* Extract observations for the ticker **SEE QUIZ** from `health_cos` and assign to the variable `health_cos_subset`


```{r eval = FALSE}
health_cos_subset  <- health_cos  %>% 
  ???(ticker == "???")

```

* Display `health_cos_subset`

```{r eval = FALSE}
health_cos_subset 

```


* In the console, type `?distinct`. Go to the help pane to see what `distinct` does
* In the console, type `?pull`.  Go to the help pane to see what `pull` does

<br> 

Run the code below

```{r eval = FALSE}
health_cos_subset  %>% 
  distinct(name) %>%  
  pull(name)

```

* Assign the output to `co_name`

```{r eval = FALSE}

???  <- health_cos_subset  %>% 
  distinct(name) %>% 
  pull(name)

```


**You can take output from your code and include it in your text.**  

* The name of the company with ticker **SEE QUIZ** is ________   

In following chuck

* Assign the company's industry group to the variable `co_industry`

```{r eval = FALSE}

co_industry  <- ???  %>% 
  ???(industry) %>% 
  ???()

```


This is outside the R chunk.  Put the r inline commands used in the blanks below.  When you knit the document the results of the commands will be displayed in your text.  

The company ??? is  a member of the ??? group.

*****

# Steps 7-11

7. Prepare the data for the plots

- start with health_cos THEN
- group_by industry THEN
- calculate the median research and development expenditure as a percent of revenue by industry
- assign the output to `df`

```{r}
df <- health_cos  %>% 
  group_by(industry)  %>%
  summarize(med_rnd_rev = median(rnd/revenue))   
```

8.  Use `glimpse` to glimpse the data for the plots

```{r}
df  %>% glimpse()
```

9. Create a static bar chart

- use `ggplot` to initialize the chart
- data is `df`
- the variable `industry` is mapped to the x-axis
    - reorder it based the value of `med_rnd_rev`
- the variable `med_rnd_rev` is mapped to the y-axis
- add a bar chart using `geom_col`
- use `scale_y_continuous` to label the y-axis with percent
- use `coord_flip()` to flip the coordinates
- use `labs` to add title, subtitle and remove x and y-axes 
- use `theme_ipsum()` from the hrbrthemes package to improve the theme


```{r}
ggplot(data = df, 
       mapping = aes(
         x = reorder(industry, med_rnd_rev ),
         y = med_rnd_rev
         )) +
  geom_col() + 
  scale_y_continuous(labels = scales::percent) +
  coord_flip() +
  labs(
    title = "Median R&D expenditures",
    subtitle = "by industry as a percent of revenue from 2011 to 2018",
    x = NULL, y = NULL) +
  theme_ipsum()
```

10. Save the previous plot to preview.png and add to the yaml chunk at the top

```{r}
ggsave(filename = "preview.png", 
       path = here::here("_posts", "2021-02-27-joining-data"))
```


11. Create an interactive bar chart using the package [echarts4r](https://echarts4r.john-coene.com/index.html)

- start with the data `df`
- use `arrange` to reorder `med_rnd_rev`
- use `e_charts` to initialize a chart 
    - the variable `industry` is mapped to the x-axis
- add a bar chart using `e_bar` with the values of `med_rnd_rev`
- use `e_flip_coords()` to flip the coordinates
- use `e_title` to add the title and the subtitle 
- use `e_legend` to remove the legends
- use `e_x_axis` to change format of labels on x-axis to percent
- use `e_y_axis` to remove labels on y-axis-
- use `e_theme` to change the theme. Find more themes [here](https://echarts4r.john-coene.com/articles/themes.html)


```{r}
df  %>% 
  arrange(med_rnd_rev)  %>%
  e_charts(
    x = industry
    )  %>% 
  e_bar(
    serie = med_rnd_rev, 
    name = "median"
    )  %>%
  e_flip_coords()  %>% 
  e_tooltip()  %>% 
  e_title(
    text = "Median industry R&D expenditures", 
    subtext = "by industry as a percent of revenue from 2011 to 2018",
    left = "center") %>% 
  e_legend(FALSE) %>% 
  e_x_axis(
    formatter = e_axis_formatter("percent", digits = 0)
    )  %>%
  e_y_axis(
    show = FALSE
  )  %>% 
  e_theme("infographic")

```

