---
title: "Project Part 2"
description: |
  Interactive and static plots of CO2 emissions from 1900 to 2019
date: 2021-11-05
output:
  distill::distill_article:
    self_contained: false
    toc: true
preview: preview.png    
---

1. Packages I will use to read in and plot the data

```{r}
library(tidyverse)
library(echarts4r)
```


2. Read the data in from part 1

```{r}
regional_co2  <- read_csv(here::here("regional_co2.csv"))

```


## Interactive graph

- Start with the data
- Group_by region so there will be a "river" for each region
- Use mutate to round CO2emissions so only 2 digits will be displayed when you hover over it. 
- Use mutate to change Year so it will be displayed as end of year instead of beginning of year 
- Use e_charts to create an e_charts object with Year on the x axis
- Use e_river to build "rivers" that contain CO2emissions by region.  The depth of each river represents the amount of emissions for each region.  
- Use e_tooltip to add a tooltip that will display based on the axis values
- Use e_title to add a title, subtitle, and link to subtitle
- Use e_theme to change the theme to roma



```{r}
regional_co2   %>%
  group_by(Region)  %>%
  mutate(CO2emissions = round(CO2emissions, 2),
         Year = paste(Year, "12", "31", sep="-"))  %>% 
  e_charts(x = Year)   %>% 
  e_river(serie = CO2emissions, legend=FALSE)  %>% 
  e_tooltip(trigger = "axis")  %>% 
  e_title(text = "Annual CO2 emissions, by world region",
          subtext = "(in billions of tonnes). Source: Our World in Data",
          sublink = "https://ourworldindata.org/co2-and-other-greenhouse-gas-emissions#global-emissions-have-not-yet-peaked",
          left = "center")  %>% 
  e_theme("roma")  
```


## Static graph

- Start with the data
- Use ggplot to create a new ggplot object.  Use aes to indicate that Year will be mapped to the x axis; CO2 emissions will be mapped to the y axis; Region will be the fill variable
- geom_area will display CO2emissions 
- scale_fill_discrete_divergingx is a function in the colorspace package. It sets the color palette to roma and selects a maximum of 12 colors for the different regions
- theme_classic sets the theme
- theme(legend.position = "bottom") puts the legend at the bottom of the plot
- labs sets the y axis label, fill = NULL indicates that the fill variable will not have the labelled Region


```{r}
regional_co2   %>% 
  ggplot(aes(x = Year, y = CO2emissions, 
             fill = Region)) +
  geom_area() +
  colorspace::scale_fill_discrete_divergingx(palette = "roma", nmax =11) +
  theme_classic() +
  theme(legend.position = "bottom") +
  labs( y = "in billions of tonnes",
       fill = NULL)
```

These plots show a steady increase in emissions since 1900.  Emissions have continued to increase.


```{r}
ggsave(filename = here::here("_posts/2021-11-05-project-part-2/preview.png"))
```
