---
title: "ggplot2 examples and exercises"
subtitle: "slides 73-108"
output: html_document
author: Thomas Lin Pederson (modified by Elizabeth Stanny)
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)
```

# Load packages

```{r load-packages, echo=FALSE}
library(pacman) #load package manager
p_load(tidyverse,
       ggrepel, gganimate, echarts4r,
       gifski, ggforce, ggthemes, 
       concaveman, hrbrthemes,  ggraph, tidygraph)
```


### `gganimate` examples and `echarts4r` code that replicates them

#### [`echarts4r`](https://www.infoworld.com/article/3607068/plot-in-r-with-echarts4r.html) is a better option for animation.

Examples using `echarts4r` follow those using `gganimate`

```{r animate-slides-74-75, eval = FALSE}
ggplot(economics) + 
  geom_line(aes(x = date, y = unemploy))

ggplot(economics) + 
  geom_line(aes(x = date, y = unemploy)) + 
  transition_reveal(along = date)
```

#### Using `echarts4r`

```{r echarts-animate-slides-74-75}
economics  %>% 
  e_charts(x = date)  %>% 
  e_line(serie = unemploy)

economics  %>%
  e_charts(x = date)  %>% 
  e_line(serie = unemploy)  %>% 
  e_animation(duration = 8000)

economics  %>%
  e_charts(x = date)  %>% 
  e_line(serie = unemploy)  %>% 
  e_line(serie = pop)  %>% 
  e_animation(duration = 8000)

```

There are many different transitions that control how data is
interpreted for animation, as well as a range of other animation
specific features

```{r animate-slides-76-77, eval = FALSE}
ggplot(mpg) + 
  geom_bar(aes(x = factor(cyl)))

ggplot(mpg) + 
  geom_bar(aes(x = factor(cyl))) + 
  labs(title = 'Number of cars in {closest_state} by number of cylinders') + 
  transition_states(states = year) + 
  enter_grow() + 
  exit_fade()
```

#### Using `echarts4r`

-   `group_by` the variable you want to transition along

-   Set `timeline = TRUE` for an animation

-   Modify an animation with
    [`e_animation`](https://echarts4r.john-coene.com/reference/e_animation.html).

-   See the article on the [timeline
    component](https://echarts4r.john-coene.com/articles/timeline.html)

```{r}
mpg  %>%
  e_charts(x = cyl)  %>% 
  e_bar(cyl)

mpg  %>% 
  group_by(year)  %>% 
  e_charts(x= cyl, timeline = TRUE)  %>% 
  e_timeline_opts(autoPlay = TRUE) %>% 
  e_bar(cyl)  %>% 
  e_timeline_serie(
    title = list(
      list(text = "1999", subtext = "Number of cars by number of cylinders"),
      list(text = "2008", subtext = "Number of cars by number of cylinders")
    )
    ) %>% 
  e_legend(FALSE)

```

#### Exercises animation

The animation below will animate between points showing cars with
different cylinders.

```{r animate-ex-1, eval = FALSE}
ggplot(mpg) + 
  geom_point(aes(x = displ, y = hwy)) + 
  ggtitle("Cars with {closest_state} cylinders") + 
  transition_states(factor(cyl))
```

gganimate uses the `group` aesthetic to match observations between
states. By default the group aesthetic is set to the same value, so
observations are matched by their position (first row of 4 cyl is
matched to first row of 5 cyl etc.). This is clearly wrong here (why?).
Add a mapping to the `group` aesthetic to ensure that points do not move
between the different states.

------------------------------------------------------------------------

In the presence of discrete aesthetic mappings (`colour` below), the
group is deduced if not given. The default behaviour of objects that
appear and disappear during the animation is to simply pop in and out of
existance. `enter_*()` and `exit_*()` functions can be used to control
this behaviour. Experiment with the different enter and exit functions
provided by gganimate below. What happens if you add multiple enter or
exit functions to the same animation?

```{r animate-ex-2, eval = FALSE}
ggplot(mpg) + 
  geom_point(aes(x = displ, y = hwy, colour = factor(cyl))) + 
  ggtitle("Cars with {closest_state} cylinders") + 
  transition_states(factor(cyl))
```

------------------------------------------------------------------------

In the animation below (as in all the other animations) the changes
happens at constant speed. How values change during an animation is
called easing and can be controlled using the `ease_aes()` function.
Read the documentation for `ease_aes()` and experiment with different
easings in the animation.

```{r animate-ex-3, eval = FALSE}

# check out package tidyr (part of tidyverse) for tidying data: https://tidyr.tidyverse.org

mpg2 <- tidyr::pivot_longer(mpg, c(cty,hwy))
ggplot(mpg2) + 
  geom_point(aes(x = displ, y = value)) + 
  ggtitle("{if (closest_state == 'cty') 'Efficiency in city' else 'Efficiency on highway'}") + 
  transition_states(name)
```

### Annotation

Text is a huge part of storytelling with your visualisation.
Historically, textual annotations has not been the best part of ggplot2
but new extensions make up for that.

Standard geom_text will often result in overlaping labels

```{r annotation-slide-80}
ggplot(mtcars, aes(x = disp, y = mpg)) + 
  geom_point() + 
  geom_text(aes(label = row.names(mtcars)))
```

ggrepel takes care of that

```{r annotation-slide-81}
ggplot(mtcars, aes(x = disp, y = mpg)) + 
  geom_point() + 
  geom_text_repel(aes(label = row.names(mtcars)))
```

If you want to highlight certain parts of your data and describe it, the
`geom_mark_*()` family of geoms have your back

```{r annotation-slide-82}
ggplot(mtcars, aes(x = disp, y = mpg)) +
  geom_point() + 
  geom_mark_ellipse(aes(filter = gear == 4,
                        label = '4 gear cars',
                        description = 'Cars with fewer gears tend to both have higher yield and lower displacement'))
```

#### Exercises - annotation

ggrepel has a tonne of settings for controlling how text labels move.
Often, though, the most effective is simply to not label everything.
There are two strategies for that: Either only use a subset of the data
for the repel layer, or setting the label to `""` for those you don't
want to plot. Try both in the plot below where you only label 10 random
points.

```{r annotation-ex-1, eval = FALSE}
mtcars2 <- mtcars
mtcars2$label <- rownames(mtcars2)
points_to_label <- sample(nrow(mtcars), 10)
ggplot(mtcars2, aes(x = disp, y = mpg)) + 
  geom_point() + 
  geom_text_repel(aes(label = points_to_label))
```

------------------------------------------------------------------------

Explore the documentation for `geom_text_repel`. Find a way to ensure
that the labels in the plot below only repels in the vertical direction

```{r annotation-ex-2, eval = FALSE}
mtcars2$label <- ""
mtcars2$label[1:10] <- rownames(mtcars2)[1:10]
ggplot(mtcars2, aes(x = disp, y = mpg)) + 
  geom_point() + 
  geom_text_repel(aes(label = label))
```

------------------------------------------------------------------------

ggforce comes with 4 different types of mark geoms. Try them all out in
the code below:

```{r annotation-ex-3}
ggplot(mtcars, aes(x = disp, y = mpg)) +
  geom_point() + 
  geom_mark_ellipse(aes(filter = gear == 4,
                        label = '4 gear cars'))
```

### Networks

ggplot2 has been focused on tabular data. Network data in any shape and
form is handled by ggraph

```{r network-slides-84-85}

graph <- create_notable('zachary') %>% 
  mutate(clique = as.factor(group_infomap()))

ggraph(graph) + 
  geom_mark_hull(aes(x, y, fill = clique)) + 
  geom_edge_link() + 
  geom_node_point(size = 2)
```

dendrograms are just a specific type of network

```{r network-slide-85}
iris_clust <- hclust(dist(iris[, 1:4]))
ggraph(iris_clust) + 
  geom_edge_bend() + 
  geom_node_point(aes(filter = leaf))
```

#### Exercies- networks

Most network plots are defined by a layout algorithm, which takes the
network structure and calculate a position for each node. The layout
algorithm is global and set in the `ggraph()`. The default `auto` layout
will inspect the network object and try to choose a sensible layout for
it (e.g. dendrogram for a hierarchical clustering as above). There is,
however no optimal layout and it is often a good idea to try out
different layouts. Try out different layouts in the graph below. See the
[the website](https://ggraph.data-imaginist.com/reference/index.html)
for an overview of the different layouts.

```{r networks-ex-1}
ggraph(graph) + 
  geom_edge_link() + 
  geom_node_point(aes(colour = clique), size = 3)
```

------------------------------------------------------------------------

There are many different ways to draw edges. Try to use
`geom_edge_parallel()` in the graph below to show the presence of
multiple edges

```{r networks-ex-2}
highschool_gr <- as_tbl_graph(highschool)
ggraph(highschool_gr) + 
  geom_edge_link() + 
  geom_node_point()
```

Faceting works in ggraph as it does in ggplot2, but you must choose to
facet by either nodes or edges. Modify the graph below to facet the
edges by the `year` variable (using `facet_edges()`)

```{r networks-ex-3}
ggraph(highschool_gr) + 
  geom_edge_fan() + 
  geom_node_point()
```

### Looks

Many people have already designed beautiful (and horrible) themes for
you. Use them as a base

```{r looks-slide-88}
p <- ggplot(mtcars, aes(mpg, wt)) +
  geom_point(aes(color = factor(carb))) +
  labs(
    x = 'Fuel efficiency (mpg)', 
    y = 'Weight (tons)',
    title = 'Seminal ggplot2 example',
    subtitle = 'A plot to show off different themes',
    caption = 'Source: It’s mtcars — everyone uses it'
  )

p + 
  scale_colour_ipsum() + 
  theme_ipsum()
```

```{r looks-slide-89}
# library(ggthemes)
p + 
  scale_colour_excel() + 
  theme_excel()
```

## Drawing anything

```{r anything-slide-92}
states <- c(
  'eaten', "eaten but said you didn\'t", 'cat took it', 'for tonight',
  'will decompose slowly'
)
pie <- data.frame(
  state = factor(states, levels = states),
  amount = c(4, 3, 1, 1.5, 6),
  stringsAsFactors = FALSE
)
ggplot(pie) + 
  geom_col(aes(x = 0, y = amount, fill = state))
```

```{r anything-slide-93}
ggplot(pie) + 
  geom_col(aes(x = 0, y = amount, fill = state)) + 
  coord_polar(theta = 'y')
```

```{r anything-slide-94}
ggplot(pie) + 
  geom_col(aes(x = 0, y = amount, fill = state)) + 
  coord_polar(theta = 'y') + 
  scale_fill_tableau(name = NULL,
                     guide = guide_legend(ncol = 2)) + 
  theme_void() + 
  theme(legend.position = 'top', 
        legend.justification = 'left')
```

```{r anything-slide-97}
ggplot(pie) + 
  geom_arc_bar(aes(x0 = 0, y0 = 0, r0 = 0, r = 1, amount = amount, fill = state), stat = 'pie') + 
  coord_fixed()
```

```{r anything-slide-98}
ggplot(pie) + 
  geom_arc_bar(aes(x0 = 0, y0 = 0, r0 = 0, r = 1, amount = amount, fill = state), stat = 'pie') + 
  coord_fixed() + 
  scale_fill_tableau(name = NULL,
                     guide = guide_legend(ncol = 2)) + 
  theme_void() + 
  theme(legend.position = 'top', 
        legend.justification = 'left')
```

```{r anything-slide-99}
ggplot(mpg) + 
  # geom_bar(aes(x = hwy), stat = 'bin')
  geom_histogram(aes(x = hwy))
```

```{r anything-slide-100}
ggplot(mpg) + 
  geom_bar(aes(x = hwy)) + 
  scale_x_binned(n.breaks = 30, guide = guide_axis(n.dodge = 2))
```
