---
title: "3: Reading and writing data"
description: |
  Practice reading and writing data, more dplyr and a plot.
date: 02-13-2021
output:
  distill::distill_article:
    self_contained: false
preview: preview.png
---

1.  Load the R packages we will use.

```{r}
library(tidyverse)
library(here)
library(janitor) # make sure you install
library(skimr)
```

2.  Download $CO_2$ emissions per capita from [Our World in Data](https://ourworldindata.org/co2/country/united-states?country=~USA#per-capita-how-much-co2-does-the-average-person-emit) into the directory for this post.

3.  Assign the location of the file to `file_csv`. The data should be in the same directory as this file

-   Read the data into R and assign it to `emissions`

```{r}
file_csv  <- here("_posts", 
                  "2021-02-13-reading-and-writing-data", 
                  "co-emissions-per-capita20220218.csv") 

emissions  <- read_csv(file_csv)

```

4.  Show the first 10 rows (observations of) `emissions`

```{r}
emissions  
```

5.  Start with `emissions` data THEN

-   use `clean_names` from the janitor package to make the names easier to work with
-   assign the output to `tidy_emissions`
-   show the first 10 rows of `tidy_emissions`

```{r}
tidy_emissions   <- emissions %>% 
  clean_names()

tidy_emissions
```

6.  Start with the `tidy_emissions` THEN

-   use `filter` to extract rows with `year == 2019` THEN
-   use `skim` to calculate the descriptive statistics

```{r}
tidy_emissions  %>% 
  filter(year == 2019)  %>% 
  skim()
```

7.  12 observations have a missing code. How are these observations different?

-   start with `tidy_emissions` then extract rows with `year == 2019` and are missing a code

```{r}
tidy_emissions  %>% 
  filter(year == 2019, is.na(code))  
```

Entities that are not countries do not have country codes.

8.  Start with tidy_emissions THEN

-   use `filter` to extract rows with year == 2019 and **without** missing codes THEN
-   use `select` to drop the `year` variable THEN
-   use `rename` to change the variable `entity` to `country`
-   assign the output to `emissions_2019`

```{r}
emissions_2019  <- tidy_emissions  %>% 
  filter(year == 2019, !is.na(code))   %>% 
  select(-year)  %>% 
  rename(country = entity)
```

9.  Which 15 countries have the highest `annual_co2_emissions_per_capita`?

- start with `emissions_2019` THEN
- use `slice_max` to extract the 15 rows with the `annual_co2_emissions_per_capita`
- assign the output to `max_15_emitters`

```{r}
max_15_emitters  <- emissions_2019  %>% 
  slice_max(annual_co2_emissions_per_capita, n = 15)
```

10.  Which 15 countries have the lowest `annual_co2_emissions_per_capita`?

- start with `emissions_2019` THEN
- use `slice_min` to extract the 15 rows with the lowest values
- assign the output to `min_15_emitters`


```{r}
min_15_emitters  <- emissions_2019  %>% 
  slice_min(annual_co2_emissions_per_capita, n = 15)
```

11. Use `bind_rows` to bind together the `max_15_emitters` and `min_15_emitters`
- assign the output to `max_min_15`


```{r}
max_min_15  <- bind_rows(max_15_emitters, min_15_emitters)
```

12. Export `max_min_15` to 3 file formats

```{r}
max_min_15  %>% write_csv("max_min_15.csv") # comma-separated values
max_min_15  %>% write_tsv("max_min_15.tsv")  # tab separated
max_min_15  %>% write_delim("max_min_15.psv", delim = "|") # pipe-separated
```

13. Read the 3 file formats into R

```{r}
max_min_15_csv <-  read_csv("max_min_15.csv") # comma-separated values
max_min_15_tsv <-  read_tsv("max_min_15.tsv")  # tab separated
max_min_15_psv <-  read_delim("max_min_15.psv", delim = "|") # pipe-separated
```

14. Use `setdiff` to check for any differences among `max_min_15_csv`, `max_min_15_tsv` and `max_min_15_psv` 

```{r eval=FALSE}
setdiff(max_min_15_csv, max_min_15_tsv, max_min_15_psv)
```

Are there any differences? 


15. Reorder `country` in `max_min_15` for plotting and assign to max_min_15_plot_data

- start with `emissions_2019` THEN
- use `mutate` to reorder `country` according to `annual_co2_emissions_per_capita`


```{r}
max_min_15_plot_data  <- max_min_15 %>%
  mutate(country = reorder(country, annual_co2_emissions_per_capita))  
```

16. Plot `max_min_15_plot_data`
 
```{r}
ggplot(data = max_min_15_plot_data, 
       mapping = aes(x= annual_co2_emissions_per_capita, y = country)) +
  geom_col() +
  labs(title = "The top 15 and bottom 15 per capita CO2 emissions",
       subtitle = "for 2019", 
       x = NULL, 
       y = NULL)  
```

17. Save the plot directory with this post

```{r}
ggsave(filename = "preview.png", 
       path = here("_posts", "2021-02-13-reading-and-writing-data"))

```

18.  Add preview.png to yaml chuck at the top of this file

```
preview: preview.png
```

