---
title: "2: Introduction to dplyr"
description: |
  Code and text for Quiz 3.
date: 02-06-2021
output:
  distill::distill_article:
    self_contained: false
---

Load the packages that we need.

```{r}
library(tidyverse)
library(readxl)
library(here)
```

Read the data into R.

```{r}
corp_tax  <- read_excel(here("corp_tax.xlsx"))
```

Let's look at **NVR** in the corp_tax tibble.

```{r}
result  <- corp_tax  %>% 
  filter(company == 'NVR')

result
```

**NVR** is in the `r result  %>% select(industry)` industry. It had profit of \$`r result  %>% select(profit)` million and tax of \$`r result  %>% select(tax)` million. Its tax rate was `r result  %>% transmute(tax_rate * 100)  %>% round(1)`%.

------------------------------------------------------------------------

Let's find the company in the **Utilities, gas and electric** industry with the highest profit

```{r}

result  <- corp_tax  %>% 
  filter(industry == 'Utilities, gas and electric')  %>% 
  slice_max(profit, n=1)
result
```

`r result  %>% select(company)` is the company in  **Utilities, gas and electric** industry with the highest profit. It had profit of \$`r result  %>% select(profit)` million and tax of \$`r result  %>% select(tax)` million. Its tax rate was `r result  %>% transmute(tax_rate * 100)  %>% round(1)`%.
